---
title: "NAPLAN in Google Trends"
author: "Neil Saunders"
date: "compiled `r Sys.time()`"
output:
  html_document:
    keep_md: yes
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.path = "../../figures/",
                      dev = c("png", "pdf"),
                      fig.width = 11, fig.height = 8.5)
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(knitr)

theme_set(theme_light())

# read data
# search 'naplan'
naplan_search_trends <- read_csv("~/Dropbox/projects/github_projects/education/naplan/data/naplan_search_trends.csv", 
    skip = 2)

# search 'naplan tests' vs 'naplan results'
naplan_search_trends_compare <- read_csv("~/Dropbox/projects/github_projects/education/naplan/data/naplan_search_trends_compare.csv", 
    skip = 2)
```

# Introduction
Visualising Google Trends data for searches related to NAPLAN testing.

# Datasets
1. CSV file with the results of a basic search for the term "naplan"
1. CSV file with the results of a comparative search for the terms "naplan tests, naplan results"

# Visualisation
## Basic search
Observations:

1. We see 2 peaks per year, corresponding to testing and release of results
1. Interest was high in 2010-11, declined, then a recent rise - related to discussion of online tests?

```{r plot-basic-search}
naplan_search_trends %>% 
  mutate(Date = as_date(paste0(Month, "-01"))) %>% 
  filter(Date > as_date("2007-12-31")) %>% 
  ggplot(aes(Date, `NAPLAN: (Australia)`)) + 
  geom_line(color = "#008acc") + 
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
    labs(y = "search volume", 
         title = "Australian Searches for term: naplan", 
         subtitle = "Source: Google Trends, 2008 - August 2017")
```

## Comparative search
Observations:

1. We see more clearly that the early peak corresponds to "tests", the later to "results"
1. In 2008: no peak for "results" inside the "tests" peak
1. In 2009: 1 peak for "results" inside the "tests" peak
1. From 2010: 2 peaks for results inside the "tests" peak, near the start and end of that peak

```{r plot-comparative-search}
naplan_search_trends_compare %>% 
  gather(search, value, -Month) %>% 
  mutate(Date = as_date(paste0(Month, "-01"))) %>% 
  filter(Date > as_date("2007-12-31")) %>% 
  ggplot(aes(Date, value)) + 
    geom_line(aes(group = search, color = search)) + 
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") + 
    scale_color_manual(values = c("darkorange", "#008acc")) + 
    labs(y = "search volume", 
         title = "Australian searches for terms: naplan tests, naplan results", 
         subtitle = "Source: Google Trends, 2008 - August 2017")
```
